---
title: "TBA2104 Predictive Analytics Tutorial/Lab 1"
author: "Chan Kong Yew"
date: "January 20, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('D:/NUS/ChadCode/TBA2104/tut01')
```


1. In this exercise, we will be using the College (College.csv) data set which can be downloaded from LumiNUS. This data set contains a number of variables for 777 different universities and colleges in the US. The variables are: 
  
    * Private : Public/private indicator
    * Apps : Number of applications received
    * Accept : Number of applicants accepted
    * Enroll : Number of new students enrolled
    * Top10perc : New students from top 10% of high school class
    * Top25perc : New students from top 25% of high school class
    * F.Undergrad : Number of full-time undergraduates
    * P.Undergrad : Number of part-time undergraduates
    * Outstate : Out-of-state tuition
    * Room.Board : Room and board costs
    * Books : Estimated book costs
    * Personal : Estimated personal spending
    * PhD : Percent of faculty with Ph.D.’s
    * Terminal : Percent of faculty with terminal degree
    * S.F.Ratio : Student/faculty ratio
    * perc.alumni : Percent of alumni who donate
    * Expend : Instructional expenditure per student
    * Grad.Rate : Graduation rate


**Note:**
At any point in the lab exercise, you can type ?func in R or RStudio to see the help on the functions  
  
    a) Download *College.csv* onto your desktop or any directory. Before reading the data into R, it can be viewed in Excel or a text editor.  
  
    b) Open RStudio and change the working directory to the path where you have placed *College.csv*  
  
    c) Use the read.csv() function to read the data into a variable college in R  
    Note: did you ensure that you do not read the header row?

```{r Part.C, echo = TRUE}

college <- read.csv('college.csv') # default header = F

```

```{r echo=FALSE}
View(college)
print(paste('college imported as: ', class(college)))
```


    d) Use the head() function to read the first 3 rows of college dataframe.  
```{r Part.D, echo=TRUE}

head(college)

```
    
    
    e) Retrieve the vector of the university names from the college dataframe.
    
```{r Part.E}

# college$X # results too long
print('college$X  results too long for page')
print(paste('show only first 10 rows: ', college[c(1:10),'X']))

```
    
    
    f) When answering (e), if there are many columns or the column names are not obvious, it is sometimes useful to be able to programmatically retrieve the column names. Try obtaining a vector of the column names of the college dataframe.  
    
```{r Part.F, echo=TRUE}

colnames(college)

```
    
  
    g) Just like how column names are used to identify the columns of a dataframe, there is also a row names for each dataframe used to identify the rows of the dataframe. Obtain a vector of the row names of the college dataframe.  
      
```{r Part.G, echo=TRUE}

rownames(college)

```
      
    h) Ideally, we want to set the row names to its corresponding university names and remove 1 the existing column of university names. What do you think are the commands to do this?  
     
```{r Part.H, echo=TRUE}

#remove rownames
rownames(college) <- NULL

#assign new name
rownames(college) <- college$X

rownames(college)

```
     
    i) What is the total number of applications received? How about the average number of applications received by a university?  
    
```{r Part.I, echo=TRUE}

#any NA?
print('any NAs in Apps col?')
which(is.na(college$Apps))

#sum up
print(paste('total applications: ', sum(college$Apps,na.rm = TRUE)))

#mean
print(paste('avg no of applications: ',mean(college$Apps,na.rm = TRUE)))

```
    
        
    j) Obtain the vector of university names that received more than 5000 applications.  
    
```{r Part.J, echo=TRUE}

college[which(college$Apps > 5000),c(which(colnames(college) == 'X'),which(colnames(college) == 'Apps'))]

```
    
        
    k) What is the average graduation rate for universities that received more than 5000 applications? Is the average graduate rate higher compared with universities that did not receive more than 5000?  
          
```{r Part.K}

C5kmore <- college[which(college$Apps > 5000),c(which(colnames(college) == 'X'),which(colnames(college) == 'Apps'),which(colnames(college) == 'Grad.Rate'))]

print(paste('avg graduation rate for universities with more than 5,000 applications: ', round(mean(C5kmore$Grad.Rate),4)))

C5kless <- college[which(college$Apps <= 5000),c(which(colnames(college) == 'X'),which(colnames(college) == 'Apps'),which(colnames(college) == 'Grad.Rate'))]

print(paste('avg graduation rate for universities with -less- than 5,000 applications: ', round(mean(C5kless$Grad.Rate),4)))

```
          
    l) Which universities have the highest/lowest number of full-time undergrads?
  
```{r Part.L, echo=TRUE}

college[which.max(college$F.Undergrad),c(which(colnames(college) == 'X'),which(colnames(college) == 'F.Undergrad'))]

college[which.min(college$F.Undergrad),c(which(colnames(college) == 'X'),which(colnames(college) == 'F.Undergrad'))]

```

